# R Primer (Syntax and Basic Functions)

## Overview of Approach
This tutorial is written as a guide for the question,"how do I do X in R?"

## Installing packages
```{r}
a = c(1,2,3)

```

```
# install packages from CRAN ----
install.packages(c("tidyverse", "devtools"))
```

### Aside: Some of my favorite packages

Install them all in one go with the code below:
```
install.packages(c("tidyverse", "devtools", "qualtRics",
                   "tidytext", "textdata", "topicmodels",
                   "wordcloud", "ggwordcloud", "lme4", "sjPlot",
                   "janitor", "esquisse", "rio",
                   "cowplot", "stringi", "digest"))

```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# load libraries ----
library(tidyverse)
set.seed(1013) # for reproducibility
```

## Getting help in R
```{r}
help("digest")
help("state")
??dplyr::mutate
??dplyr
```

# Quick note on case-sensitivity

```{r}
a = "apples"
A = "bananas"
print(a)
print(A)
a == A
```

## Variable Assignment

Helpful article, [RE: <- and = for assignment](https://www.roelpeters.be/the-difference-between-arrow-and-equals-assignment-in-r/)

tldr, to equate R with variable assignment in other programming languages, the = symbol was implemented.
```{r}
b = "bot"
c <- "bot"
b
c

`<-`(x, 3) # Prefix/Polish notation of arrow shortcut
x <- 3 # Leftwards assignment
3 -> x # Rightwards assignment
x = 3
x <<- 3 #  it will store the variable in the global environment (.GlobalEnv). 
```

## Assign values into variables

```{r, message=FALSE, warning=FALSE}
this_is_double <- 1
this_is_double2 = 1.0
this_is_logical <- TRUE
this_is_date <- as.Date("2020-10-13")
this_is_list <- list(x = rnorm(100), y = rnorm(10))
this_is_dataframe <- data.frame(x = rnorm(100), y = rnorm(100))
this_is_tibble <- tibble::tibble(x = rnorm(100), y = rnorm(100))
this_is_vector <- c(1,2,3)
```


## Check underlying types
```{r}
class(this_is_double)
class(this_is_double2)
class(this_is_logical)
class(this_is_date)
class(this_is_list)
class(this_is_dataframe)
class(this_is_tibble)
class(this_is_vector)
```

```{r, message=FALSE, warning=FALSE}
# generate a sequence ----
seq_1_100 <- seq(1, 100, by=1)
seq_0_1 <- seq(0, 1, by=0.1)
seq_0_1

# sample from vector of values -----
list_to_sample <- c("A", "1", "2", "B")
sampled_value = sample(list_to_sample, size=1)
sampled_value
sampled_values_replace = sample(list_to_sample, size=10, replace=TRUE)
sampled_values_replace
```

## File operations

```{r, message=FALSE, warning=FALSE}
# list files
all_csv_files = list.files("../data", 
                       pattern="*.csv", 
                       recursive = T, 
                       full.names = T)
all_csv_files

all_files = list.files("../data", 
                       pattern="*", 
                       recursive = T, 
                       full.names = T)
all_files

# check if directory exists
dir.exists("../data")
dir.exists("data")

# create directory
dir.create("test")

# create file if doesn't exist
file <- "test/dataframe.R"
if (file.exists(file)) {
  cat("The file already exists")
} else {
  file.create(file)
  print(paste0("File created: ",file))
}

# remove directory
unlink("test", recursive=TRUE)

```

## Reading and writing files

```{r}
library(readxl)
library(readr)
library(haven)

# write files of
df_for_examples = tibble(x = rnorm(100), y = rnorm(100), study="demo_rnorm_in_r")
```

```{r}
library(readr)
write_csv(df_for_examples, "../data/demo_fileread.csv")
write_delim(df_for_examples,"../data/demo_fileread.txt",delim="\t")
write_sav(df_for_examples, "../data/demo_fileread.sav")
write_sas(df_for_examples, "../data/demo_fileread.sas")
```

## Export multiple dataframes into a single Excel workbook
```{r}
library(rio)
df_1 = data.frame(x= rnorm(10), dataset=1)
df_2 = data.frame(x= rnorm(10), dataset=2)
df_3 = data.frame(x= rnorm(10), dataset=3)

export(format="xlsx", file="../output/OneExcelWorkbookWith3Sheet.xlsx", x=list(df1=df_1, df2=df_2, df3=df_3))
```

## Sanitize column names

```{r}
library(janitor)

# create demo data
df_janitor_before = tibble::tibble(`A long column Name with Mixed case`=rnorm(20),
                    `COLUMN_NAME`=rnorm(20),
                    `GROUP_$$`=rnorm(20))
knitr::kable(df_janitor_before)

df_janitor_after = df_janitor_before %>%
  janitor::clean_names()
knitr::kable(df_janitor_after)

```

```{r}
# load data ----

# plaintext data files
library(readr)

# csv
dataset_csv <- read_csv("../data/demo_fileread.csv", na = "empty")

# tab-delimited file
dataset_tab <- read_delim("../data/demo_fileread.txt", delim = "\t", 
                      escape_double = FALSE, 
                      trim_ws = TRUE)

# Excel file
library(readxl)
dataset_excel <- read_excel("../data/demo_fileread.xlsx", sheet = "Sheet1")

# read SPSS and SAS data file
library(haven)
dataset_sav <- read_sav("../data/demo_fileread.sav")
#dataset_sas <- read_sav("data/demo_fileread.sas")
```


```{r}
# dataset operations -----

# stack two or more datasets
df1 = data.frame(x= rnorm(10), dataset=1)
df2 = data.frame(x= rnorm(10), dataset=2)
df3 = data.frame(x= rnorm(10), dataset=3)
df_all = bind_rows(df1, df2, df3)

# join datasets by common identifier
df_link <- data.frame(dataset=c(1,2,3,4), description=c("mock1", "mock2", "mock3","mock4"))
df_all_link_innerjoin <- df_all %>% inner_join(df_link)
df_all_link_fulljoin <- df_all %>% full_join(df_link) # end up with 1 extra due to entry for dataset 4

```

## Binning data

```{r}

n = 50
data_to_bin = data.frame(x = rnorm(n=n,mean=500))
data_binned_label = data_to_bin %>% mutate(x_bin = cut(x, breaks=c(0,300,400,500,600,Inf)))
data_binned_num = data_to_bin %>% mutate(x_bin = cut(x, breaks=c(0,300,600,800), labels=FALSE))
data_binned_ord = data_to_bin %>% mutate(x_bin = cut(x, ordered_result=T, breaks=c(0,300,400,500,600,Inf)))

knitr::kable(head(data_binned_label))
knitr::kable(head(data_binned_num))

```


## String manipulations
```{r}

## Search for data in data
grep("APP", c("APPLES", "APP", "APPLICATION"))
grep("APP", c("APPLES", "APP", "APPLICATION"), value=T)

c("APPLES", "APP", "APPLICATION")[grepl("APP", c("APPLES", "APP", "APPLICATION"))]

## Replace data in data
gsub("APPLES","BANANAS",c("APPLES", "BANANAS"))
gsub("ERR","ERROR",c("ERR", "ERRROR"))

gsub("stimuli:","",c("stimuli:word", "stimuli:word2"))
```

```{r}
# stringi stuff
library(stringi)

stri_count_fixed("ACAACATGAACGGGTACACACTG", "ACA", overlap=TRUE)

x <- c("spam", "")
stri_length(x)

stri_dup(letters[1:26], 1:26)

words <- list(c("spam", "bacon", "sausage", "spam"), c("eggs", "spam"))
stri_join_list(words, sep=":")
```

## Generate Random data

```{r}
x = rnorm(n=500,mean=300,sd=1)
hist(x)
```

## Simple stats

```{r}
mean(x)
sd(x)
min(x)
max(x)
quantile(x,c(0.6))
quantile(x,c(0.95))
quantile(x,c(0.99))

quantile(x,c(0.01,0.99))
```

## Data quality reports

```{r}
## Data quality inspection
knitr::kable(skimr::skim(x)) # one variable 
knitr::kable(skimr::skim(dataset_csv)) # or a whole dataset!
dqr = skimr::skim(dataset_csv) %>%
  mutate(dataset = "fake")# save to a variable to later export as csv
knitr::kable(dqr)
```

## Data integrity

MD5 and SHA256 hashes of data are helpful when making snapshots of data for long-term storage and curation. It ensures that the receiver can verify this hash upon opening the dataset and knowing that no data was lost or corrupted in transmission.

```{r}
library(digest)
digest("a ", algo="md5")
digest(dqr, algo="md5")
digest("a", algo="sha256")
```
