# Data Wrangling - Special considerations with survey data

## Overview

In this example, we will work with the `qualtRics` package to read in Qualtrics exports, and apply minimal pre-processing.

```{r}
install.packages("kableExtra")
```

## Load data

Read in survey and split StartDate into two columns for easy filtering (we don't need pilot data)

```{r}
library(qualtRics)
library(tidyverse)

# set filename of survey file
fn = "../data/demo_qualtrics_export.csv"

# filter out pilot data
PILOT_START_DATE = as.Date("2022-06-21")
qualtrics_df <- read_survey(fn) %>%
  separate(StartDate, into=c("dt_date", "dt_time"), sep=" ", remove=F) %>%
  mutate(is_pilot_data = dt_date < PILOT_START_DATE) %>%
  filter(!is_pilot_data) %>%
  mutate(participant_id = as.character(ResponseId))
```

## Split demographics into seperate dataset
```{r}
df_demo = qualtrics_df %>%
select(ResponseId, contains("demo"), -contains("timer"))

knitr::kable(df_demo) %>% kableExtra::kable_paper(.)
knitr::kable(df_demo) %>% kableExtra::kable_material(.)

knitr::kable(df_demo) %>% kableExtra::kable_material_dark(.)
```

## Split out survey metadata
```{r}
df_metadata = qualtrics_df %>%
select(ResponseId, contains("Date"), contains("duration"), contains("progress"))

knitr::kable(df_metadata) %>% kableExtra::kable_paper(.)
```

## Split timer data into seperate dataset
```{r}
df_timings = qualtrics_df %>%
select(ResponseId, contains("timer"))

knitr::kable(df_timings) %>% kableExtra::kable_paper(.)

```

## Score TIPI data

### Create reverse-scoring function

```{r}
reverse_code <- function(df, ques_label) {
  out_df = df %>%
    filter(tipi_question == ques_label) %>%
    mutate(response_n = recode(response, 
                               `Disagree strongly` = 7,
                               `Disagree moderately` = 6, 
                               `Disagree a little` = 5,
                               `Neither agree nor disagree` = 4,
                               `Agree a little` = 3, 
                               `Agree moderately` = 2,
                               `Agree strongly` = 1,
                               .default=-999))
  return(out_df)
}
```

### Seperate TIPI data from main `qualtrics_df`

```{r}
# create TIPI dataset -----
tipi_df = qualtrics_df %>%
  select(ResponseId, contains("TIPI"), -contains("_Timer"))
```

### Transform TIPI from wide to long
```{r}
tipi_long = tipi_df %>%
  pivot_longer(cols=TIPI_1:TIPI_10, names_to="tipi_question", values_to="response")
```

### Recode values
```{r}
tipi_long_r = tipi_long %>%
  mutate(response_n = recode(response, 
                             `Disagree strongly` = 1,
                             `Disagree moderately` = 2, 
                             `Disagree a little` = 3,
                             `Neither agree nor disagree` = 4,
                             `Agree a little` = 5, 
                             `Agree moderately` = 6,
                             `Agree strongly` = 7,
                             .default=-999))

```

### Apply Reverse Coding

NOTE: Extraversion: 1, 6R; Agreeableness: 2R, 7; Conscientiousness; 3, 8R; Emotional Stability: 4R, 9; Openness to Experiences: 5, 10R

```{r}
tipi_long_recode2 = reverse_code(tipi_long_r, "TIPI_2")
tipi_long_recode4 = reverse_code(tipi_long_r, "TIPI_4")
tipi_long_recode6 = reverse_code(tipi_long_r, "TIPI_6")
tipi_long_recode8 = reverse_code(tipi_long_r, "TIPI_8")
tipi_long_recode10 = reverse_code(tipi_long_r, "TIPI_10")

tipi_long_nonrecoded = tipi_long_r %>%
  filter(tipi_question %in% c("TIPI_1","TIPI_3","TIPI_5","TIPI_7","TIPI_9"))

tipi_long_recode_all = bind_rows(tipi_long_nonrecoded,
                                 tipi_long_recode2,
                                 tipi_long_recode4,
                                 tipi_long_recode6,
                                 tipi_long_recode8,
                                 tipi_long_recode10)
```

### Finally transform to wide format

```{r}
tipi_wide_recoded = tipi_long_recode_all %>%
  select(-response) %>%
  pivot_wider(names_from="tipi_question", values_from="response_n") %>%
  rowwise() %>%
  mutate(tipi_extraversion = mean(c(TIPI_1, TIPI_6), na.rm=T)) %>%
  mutate(tipi_agreeableness = mean(c(TIPI_2, TIPI_7), na.rm=T),
         tipi_conscientiousness = mean(c(TIPI_3, TIPI_8), na.rm=T),
         tipi_emostability = mean(c(TIPI_4, TIPI_9), na.rm=T),
         tipi_openexperience = mean(c(TIPI_5, TIPI_10), na.rm=T))

# notice the parameter `ignore.case` = F, we want this since
# we have both TIPI_* and tipi_* variables in our dataset.
knitr::kable(tipi_wide_recoded %>% 
select(contains("TIPI_", ignore.case = F))) %>% kableExtra::kable_paper(.)

knitr::kable(tipi_wide_recoded %>% 
select(contains("tipi_", ignore.case = F))) %>% kableExtra::kable_paper(.)
```
