# Data Wrangling and Visualization of Big Data

## Overview

For this example, we are working with the Google Community Mobility Reports dataset.

"These Community Mobility Reports aim to provide insights into what has changed in response to policies aimed at combating COVID-19. The reports chart movement trends over time by geography, across different categories of places such as retail and recreation, groceries and pharmacies, parks, transit stations, workplaces, and residential."

[Learn More](https://www.google.com/covid19/mobility/)

## Load libs/packages

```{r}

library(readr)
library(tidyverse)

```

```{r}
is_workshop = T
```

## Read in data
```{r}

if(!is_workshop) {
  # read in the Google Mobility dataset
  google_mobility <- read_csv("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv")
  # ~ 700MB AND 7MIN (LAST UPDATED: 07/07/2022)
} else {
  load("gmd.RData")
}

```

## Inspect dataset features

```{r}
nrow(google_mobility)
ncol(google_mobility)
```

## Create lookup tables

R, as well as some packages, have built in datasets. They can be called with the data() function (e.g., data(state)). We will leverage a built in dataset with state metadata.

```{r}
data(state)
state_lookup = tibble(sub_region_1 = state.name,
                      us_division = state.division,
                      us_region = state.region,
                      state_land_area = state.area)

# for more state data, look at the following data object (data from 1970s)
state.x77
```

## Pre-process data

  - filter data
  - add meta-date features (month, year, day, etc)
  
```{r}

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

# filter for US
google_mobility_US <- google_mobility %>%
  filter(country_region == "United States") %>%
  # filter(country_region %in% c("United States", "United Arab Emirates")) %>%
  mutate(dt_date = date) %>%
  mutate(dt_month = lubridate::month(dt_date),
         dt_month_ = lubridate::month(dt_date, label=T),
         dt_year= lubridate::year(dt_date),
         dt_day = lubridate::day(dt_date),
         dt_wkday_l = lubridate::wday(dt_date, label=T),
         dt_wkday_n = lubridate::wday(dt_date)) %>%
  mutate(dt_weekend = ifelse(dt_wkday_l %in% c("Sat", "Sun"), TRUE, FALSE))

```

## Joining datasets

  - Merge Google Mobility dataset with R data for US states

```{r}
google_mobility_US_w_lu <- google_mobility_US %>%
  inner_join(state_lookup)
```

## Quick descriptives

  - single line of code descriptives across a dataset

```{r}

dqr_gm = skimr::skim(google_mobility_US)
head(dqr_gm)
knitr::kable(dqr_gm) %>% kableExtra::kable_classic()
```

## Filter data

  - filter with specific entries
  - filter for a collection of entries

```{r}

# filter with specific entries
google_mobility_US_NY <- google_mobility_US %>%
  filter(sub_region_1 == "New York")

google_mobility_US_FL <- google_mobility_US %>%
  filter(sub_region_1 == "Florida")

google_mobility_US_neFL <- google_mobility_US %>%
  filter(sub_region_1 != "Florida")

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

# filter to a 'vector' of entries
google_mobility_US_targeted <- google_mobility_US %>%
  filter(sub_region_1 %in% c("California", "New York", "Florida", "Alaska", "Texas"))

```

## Compare states
```{r}
library(ggstatsplot)
ggbetweenstats(
  data  = google_mobility_US_targeted,
  x     = sub_region_1,
  y     = retail_and_recreation_percent_change_from_baseline,
  title = "Distribution of retail change across states"
)
```

```{r}
grouped_ggbetweenstats(
  data  = google_mobility_US_targeted,
  x     = sub_region_1,
  y     = retail_and_recreation_percent_change_from_baseline,
  grouping.var     = dt_weekend,
  outlier.tagging  = TRUE,
  #outlier.label    = title,
  outlier.coef     = 2,
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
  p.adjust.method  = "bonferroni",
  palette          = "default_jama",
  package          = "ggsci",
  plotgrid.args    = list(nrow = 1),
  annotation.args  = list(title = "Differences in retail change by state for different days of week")
)
```

## Compute aggregates

```{r}

# base R approach
table(google_mobility_US$sub_region_1)

# number of records by state
records_per_state = google_mobility_US %>%
  group_by(sub_region_1) %>%
  summarise(n_records = n()) %>%
  arrange(sub_region_1)
records_per_state

```

```{r}

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

# group data by weekday
metrics_month_weekend_US <- google_mobility_US %>%
  group_by(sub_region_1, dt_year, dt_month_) %>%
  summarise(avg_change_parks = mean(parks_percent_change_from_baseline, na.rm=T),
            avg_change_retail_rec = mean(retail_and_recreation_percent_change_from_baseline, na.rm=T),
            avg_change_workplaces = mean(workplaces_percent_change_from_baseline, na.rm=T)) %>%
  inner_join(state_lookup)
```

## Visualize the data

```{r}

# plot summary data across all states
allUS_retail = ggplot(metrics_month_weekend_US %>% 
                        filter(us_region %in% c("South", "West")), aes(sub_region_1, avg_change_retail_rec)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=90)) +
  facet_wrap(dt_year~.) +
  labs(x="Month", y="Retail (% Change)")

allUS_parks = ggplot(metrics_month_weekend_US %>% filter(us_region %in% c("South", "West")), aes(sub_region_1, avg_change_parks)) +
  geom_boxplot() +
  theme_minimal() +
  facet_wrap(dt_year~.) +
  theme(axis.text.x=element_text(angle=90)) +
  labs(x="Month", y="Parks (% Change)")

allUS_retail
allUS_parks
```

```{r}
jitter_retail <- ggplot(metrics_month_weekend_US, aes(dt_month_, 
                                     avg_change_retail_rec, 
                                     group=us_region,
                                     color=us_region)) +
  #geom_point() +
  geom_jitter() +
  theme_minimal() +
  labs(x="Month", y="Retail & Recreation (% Change)")

smooth_retail = ggplot(metrics_month_weekend_US, aes(dt_month_, 
                                     avg_change_retail_rec, 
                                     group=us_region,
                                     color=us_region)) +
  #geom_point() +
  #geom_jitter() +
  geom_smooth() +
  theme_minimal() +
  labs(x="Month", y="Retail & Recreation (% Change)")

smooth_parks = ggplot(metrics_month_weekend_US, aes(dt_month_, 
                                     avg_change_parks, 
                                     group=us_region,
                                     color=us_region)) +
  #geom_point() +
  #geom_jitter() +
  geom_smooth() +
  theme_minimal() +
  labs(x="Month", y="Parks (% Change)")

jitter_retail
smooth_retail
smooth_parks
```
[Click here to learn more about geom_smooth](https://ggplot2.tidyverse.org/reference/geom_smooth.html)

## Combining multiple plots
```{r}
library(cowplot)

# create a plot of weekly trends
early_covid_retail = ggplot(google_mobility_US_targeted %>%
         filter(dt_month %in% c(1,2,3,4,5)) %>%
         filter(dt_year == 2020), aes(dt_date, 
                                    retail_and_recreation_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  geom_jitter() +
  theme_minimal() +
  facet_wrap(.~sub_region_1) +
  labs(x="Month", y="Retal & Rec (% Change)", title="2020")

early_covid_parks = ggplot(google_mobility_US_targeted %>%
         filter(dt_month %in% c(1,2,3,4,5)) %>%
         filter(dt_year == 2020), aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  geom_jitter() +
  theme_minimal() +
  facet_wrap(.~sub_region_1) +
  labs(x="Month", y="Parks (% Change)", title="2020")

nxtyear_covid_retail = ggplot(google_mobility_US_targeted %>%
         filter(dt_month %in% c(1,2,3,4,5)) %>%
         filter(dt_year == 2021), aes(dt_date, 
                                    retail_and_recreation_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  geom_jitter() +
  theme_minimal() +
  facet_wrap(.~sub_region_1) +
  labs(x="Month", y="Retal & Rec (% Change)", title="2021")

```

```{r}
cowplot::plot_grid(early_covid_retail, early_covid_parks, ncol=1)
cowplot::plot_grid(early_covid_retail, nxtyear_covid_retail, ncol=1)
cowplot::plot_grid(allUS_retail, allUS_parks, ncol=1)
cowplot::plot_grid(smooth_retail, smooth_parks, ncol=1)
```

## Filter dataset for next plots
```{r}
# Variable names -----
#   retail_and_recreation_percent_change_from_baseline = col_double(),
#   grocery_and_pharmacy_percent_change_from_baseline = col_double(),
#   parks_percent_change_from_baseline = col_double(),
#   transit_stations_percent_change_from_baseline = col_double(),
#   workplaces_percent_change_from_baseline = col_double(),
#   residential_percent_change_from_baseline = col_double()


early_covid_period_data = google_mobility_US_w_lu %>%
         filter(dt_month %in% c(1,2,3,4,5,6)) %>%
         filter(dt_year == 2020) %>%
         filter(sub_region_1 %in% c("New York", "California", "Florida", "Wyoming", "Texas", "Arizona","Montana", "South Dakota", "Nevada"))

```

## Styling your plots (theme_*)

```{r}
ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  theme_void() +
  geom_jitter() +
  facet_wrap(.~sub_region_1) +
  theme(legend.position = "none")

ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=sub_region_1)) +
  geom_jitter() +
  facet_wrap(.~sub_region_1)

ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  theme_bw() +
  geom_jitter() +
  facet_wrap(.~sub_region_1)

ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=sub_region_1)) +
  #geom_point() +
  theme_dark() +
  geom_jitter() +
  facet_wrap(.~sub_region_1)
```

## Production plots

```{r}
ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=us_region)) +
  #geom_point() +
  theme_bw() +
  geom_jitter() +
  facet_wrap(.~sub_region_1) +
  labs(x= "Date", y="Retail & Recreation (% Change)")
```

```{r}
ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=us_region)) +
  #geom_point() +
  geom_jitter() +
  facet_wrap(.~sub_region_1) +
  labs(x= "Date", y="Retail & Recreation (% Change)") +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle=90))
```

```{r}
fig1 <- ggplot(early_covid_period_data, aes(dt_date, 
                                    retail_and_recreation_percent_change_from_baseline,
                                    color=us_region)) +
  #geom_point() +
  #geom_jitter() +
  geom_smooth() +
  facet_wrap(.~sub_region_1) +
  labs(x= "Date", y="Retail & Recreation (% Change)",
       #color="State",
       tag = "Figure 1",
       title = "Google Mobility: Change over Time",
       subtitle = "i.e., Retail & Recreation, early COVID-19 period (01/20 - 06/20)",
       caption = "Google Mobility: Retail & Recreation (% Change) over early COVID period (01/20 - 06/20)") +
  theme_bw(base_size = 12, base_family = "Arial")

fig2 <- ggplot(early_covid_period_data, aes(dt_date, 
                                    parks_percent_change_from_baseline,
                                    color=us_region)) +
  #geom_point() +
  #geom_jitter() +
  geom_smooth() +
  facet_wrap(.~sub_region_1) +
  labs(x= "Date", y="Parks (% Change)",
       #color="State",
       tag = "Figure 2",
       title = "Google Mobility: Change over Time",
       subtitle = "i.e., Parks, early COVID-19 period (01/20 - 06/20)",
       caption = "Google Mobility: Parks (% Change) over early COVID period (01/20 - 06/20)") +
  theme_bw(base_size = 12, base_family = "Arial") +
  scale_color_viridis_d()
fig1
fig2
```

## Evaluate regions of most change
```{r}

largest_changes_parks_long <- google_mobility_US %>%
  group_by(sub_region_1, dt_year) %>%
  summarise(avg_change_parks = mean(parks_percent_change_from_baseline, na.rm=T)) %>%
  arrange(sub_region_1)

largest_changes_parks_wide <- largest_changes_parks_long %>%
  #pivot_wider(id_cols=c())
  pivot_wider(names_from = dt_year, values_from = avg_change_parks) %>%
  mutate(change_2022_2020 = `2022` - `2020`)

top10_changes <- largest_changes_parks_wide %>%
  arrange(-change_2022_2020) %>%
  head(n=4)

bottom10_changes <- largest_changes_parks_wide %>%
  arrange(change_2022_2020) %>%
  head(n=4)

largest_changes_parks_long
largest_changes_parks_wide

knitr::kable(top10_changes) %>% kableExtra::kable_paper()
knitr::kable(bottom10_changes) %>% kableExtra::kable_paper()
```
