# Working with APIs

## Overview

APIs make the world go round. From airlines, to your groceries, APIs are charged with securely and efficiently shipping data back and forth between businesses and their consumers.

This example will demonstrate how to interact with a set of APIs that cover a range of possible use cases (e.g., paginated and non-paginated returns, authentication).

## Public APIs as a Learning Tool
Below is a list of public APIs you may consider for your learning:

  - [https://github.com/public-apis/public-apis](https://github.com/public-apis/public-apis)
  - [Randomly Generated Data](https://random-data-api.com/documentation)

## Get 10 records of faux leaked user records
https://random-data-api.com/api/internet_stuff/random_internet_stuff

### Custom function to call API and return data frame
```{r}
call_api <- function(api) {
return(as.data.frame(jsonlite::read_json(api)))
}
```

```{r}
library(tidyverse)

n_records = 10
results = list()
for(i in 1:n_records) {
  results[[i]] = call_api("https://random-data-api.com/api/internet_stuff/random_internet_stuff")
}
results_df = bind_rows(results)
```

```{r}
library(knitr)
library(kableExtra)
knitr::kable(results_df) %>% kableExtra::kable_minimal(.)
```
