# Working with APIs

## Overview

APIs make the world go round. From airlines, to your groceries, APIs are charged with securely and efficiently shipping data back and forth between businesses and their consumers.

This example will demonstrate how to interact with a set of APIs that cover a range of possible use cases (e.g., paginated and non-paginated returns, authentication).

:::attention
## Be Mindful of Limits
When connecting to APIs for data, be mindful of the frequency of requests you make. With overuse, depending on the API, you may be throttled, blocked, or limited for a period of time. Check API documentation for posted limits (e.g., 1000 requests per month).
:::

## Public APIs as a Learning Tool
Below is a list of public APIs you may consider for your learning:

  - [https://github.com/public-apis/public-apis](https://github.com/public-apis/public-apis)
  - [Randomly Generated Data](https://random-data-api.com/documentation)

## The Data

For this learning exercise, we will work with an API that generates random data. We will get 10 records of faux user records.

https://random-data-api.com/api/internet_stuff/random_internet_stuff

### Custom function to call API and return data frame
```{r}
call_api <- function(api) {
return(as.data.frame(jsonlite::read_json(api)))
}
```

```{r}
library(tidyverse)

n_records = 10
results = list()
for(i in 1:n_records) {
  results[[i]] = call_api("https://random-data-api.com/api/internet_stuff/random_internet_stuff")
}
results_df = bind_rows(results)
```

```{r}
library(knitr)
library(kableExtra)
knitr::kable(results_df, table.attr = "style = \"color: black;\"") %>% kableExtra::kable_minimal(.)
```
