# Regression, and Multilevel Modeling

## Overview
This example uses data from a reaction time experiment where observers searched for a green circle amongst rectangles, and were asked to report if a line inside the green circle was vertical or horizontal. On some trials, a red rectangle was present, i.e, a distractor. The experiment was conducted in two ways: (1) a random ordering of distractor present and absent trials, and (2) a blocked ordering of distractor present and absent trials. With blocked administration, presumably participants had time to habituate to the distractor and would have a smaller response time decrement as compared with the random presentation condition. 

In this example analysis, we will first take a simple approach, and conduct a linear regression, on participant block-level means of response time. Thereafter we will apply multilevel modeling to better capture the nesting and variability inherent in the trial-level data.

## Load libs

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(lme4)
```

## Load data
```{r}
exp_df <- read_csv("data/exp1.csv")
```

```{r}
# show data structure
head(exp_df)
```
---

## Pre-process data
```{r}
# filter
exp_pp <- exp_df %>%
  filter(!is.na(PARTICIPANT))
```

### Create additional columns, remove columns
```{r}

exp_pp_ = exp_pp %>%
  mutate(is_practice = ifelse(practice == "yes", 1, 0)) %>%
  mutate(TRIAL = count_trial_sequence) %>%
  select(-count_trial_sequence, -response_time, -practice) %>%
  select(datetime, PARTICIPANT, BLOCK, TRIAL, is_practice, correct, RT, everything()) %>%
  mutate(EXP_BLOCK_RANDOM = ifelse(isRandom == "blocked", 0, 1))
```

---

### Get unique metadata
```{r}
metadata = exp_pp_  %>%
  select(datetime, contains("DT_"), PARTICIPANT, contains("EXP_")) %>%
  distinct()
```

---

## Produce simple aggregates

```{r}
# overall means and record count by participant
part_means = exp_pp_ %>%
  group_by(PARTICIPANT, datetime) %>%
  summarise(mean_rt = mean(RT, na.rm=T),
            sd_rt = sd(RT, na.rm=T),
            q10_rt = quantile(RT, probs=0.1, na.rm=T),
            q90_rt = quantile(RT, probs=0.9, na.rm=T),
            n_correct = sum(correct),
            n = n()) %>%
  inner_join(metadata)

# overall means by block
part_means_byblock = exp_pp_ %>%
  group_by(PARTICIPANT, datetime, BLOCK, is_practice, distractor) %>%
  summarise(mean_rt = mean(RT, na.rm=T),
            sd_rt = sd(RT, na.rm=T),
            q10_rt = quantile(RT, probs=0.1, na.rm=T),
            q90_rt = quantile(RT, probs=0.9, na.rm=T),
            n_correct = sum(correct),
            n = n()) %>%
  inner_join(metadata)
```
### Mean center the reaction times
```{r}
exp_m = exp_pp_ %>%
  inner_join(part_means) %>%
  mutate(imean_c_RT = RT - mean_rt)
```
---

## Data visualization

```{r}
ggplot(part_means_byblock %>% filter(!is_practice), aes(BLOCK, mean_rt, group=BLOCK, color=as.factor(EXP_BLOCK_RANDOM))) + 
  geom_boxplot() +
  facet_grid(. ~ EXP_BLOCK_RANDOM) +
  theme_bw() +
  labs(x= "Experiment Block", y = "Mean Response Time (ms)")
```

---

## Predictive Modeling

## Run simple regression on block means
```{r}
fit = lm(mean_rt ~ n_correct + distractor, data=part_means_byblock)
summary(fit)
```

### Show resulting model equation
```{r}
# show model equation
library(equatiomatic)
extract_eq(fit, use_coefs=TRUE, wrap=TRUE)
```

## Run as series of incremental multilevel model
```{r}
# run models, starting with unconditional means model
fitm_0  = lme4::lmer(RT ~ 1 + (1|PARTICIPANT),
                      data=exp_pp_)
fitm_1  = lme4::lmer(RT ~ TRIAL + (1|PARTICIPANT),
                      data=exp_pp_)
fitm_2  = lme4::lmer(RT ~ TRIAL + is_practice + (1|PARTICIPANT),
                      data=exp_pp_)
fitm_3  = lme4::lmer(RT ~ TRIAL + is_practice + correct + (1|PARTICIPANT),
                      data=exp_pp_)
fitm_4  = lme4::lmer(RT ~ TRIAL + is_practice + correct +
                       distractor + 
                       (1|PARTICIPANT),
                      data=exp_pp_)
fitm_5  = lme4::lmer(RT ~ TRIAL + is_practice + correct +
                       distractor + 
                       display_size + 
                       EXP_BLOCK_RANDOM +
                       (1|PARTICIPANT),
                      data=exp_pp_)
```

```{r}
extract_eq(fitm_5, use_coefs=TRUE, wrap=TRUE)
```

### Compare models
```{r}
# compare models
anova(fitm_0, fitm_1, fitm_2, fitm_3, fitm_4, fitm_5)
```

## Create publication-ready table of results
```{r}
sjPlot::tab_model(fitm_0, fitm_5)
```

## Create publication-ready figure of results
```{r}
sjPlot::plot_model(fitm_5, type="pred")
```

